Bipartite Fidelity and Loschmidt Echo of Bosonic Conformal Interface

### 2017-05-11 general_formulation.tex
[draft] appendix done; analytic results done; discussion draft 1.

\todo[inline]{Mao: I put $\partial_t$ on the top in my notes. The 'DN' 'ND' convention seems to be wrong here. I'll think about it.}
\begin{eqnarray}\begin{aligned}
\label{eq:def_M}
\begin{pmatrix}
\partial_t\phi^1\\
\partial_x\phi^1
\end{pmatrix}
=M\begin{pmatrix}
\partial_t\phi^2\\
\partial_x\phi^2
\end{pmatrix}
\end{aligned}\end{eqnarray}
where derivative should be understood in the appropriate left and right limits, for example $\partial_x \phi^1$ is evaluated at $ x = 0^-$. As argued, the momentum component of stress tensor is continuous across the interface. As a consequence $M$ is an element of the Lorentz group $O(1,1)$ and can be parameterized as
\begin{eqnarray}\begin{aligned}
M_1(\theta)=\pm
\begin{bmatrix}
\lambda^{-1} & 0 \\
0 & \lambda
\end{bmatrix}\quad
M_2(\theta)=\pm
\begin{bmatrix}
0 & \lambda  \\
\lambda^{-1} & 0 
\end{bmatrix}
\end{aligned}\end{eqnarray}
where $\lambda=\tan\theta$ for $\theta\in\left[-\frac{\pi}{2},\frac{\pi}{2}\right]$.

Several special choice of $\theta$ need to be noted. 
\begin{enumerate}
\item $\theta=0,\pm \frac{\pi}{2}$. In this case, $\lambda$ (or $\lambda^{-1}$) appears to be singular and the fields on either side of the defect cannot transmit to the other side, so the interface reduces to individual boundary conditions for the boson on the left and right half planes: they are combination of Dirichlet and Neumann boundary conditions. For example, $\lambda = 0$ for $M_1$ implies $\partial_x\phi^1 = \partial_t\phi^2 =0$, which means that Dirichlet boundary condition are imposed on the right and Neumann boundary condition on the left. Hereafter we shall denote this combination as 'DN'. Similarly $M_1(\pm\frac{\pi}{2}),M_2(0),M_2(\pm \frac{\pi}{2})$ correspond to DN, DD, NN respectively. 
\item $\theta = \pm \frac{\pi}{4}$. In this case, $M(\theta)$ characterizes a perfectly transmitting defect. For example, there is effectively no defect in the case $M( \frac{\pi}{4})$. For the other three cases, despite picking up a phase, the two counter propagating modes are still fully transmitted cross the interface. 
\end{enumerate}

The physical significance of $\theta$ will be manifest in the scattering process described below. We rewrite the Eq.~\eqref{eq:def_M} in the coordinates $t\pm x$ and $\partial_{\pm} = \partial_{t} \pm \partial_x $ derivative to extract the left and right going modes. For example, $\partial_{-} \phi^2$ will be a function $t - x$ only and hence represent a right going mode on the right half plane. It is then one of the scattering products that leave the interface. On the other hand $\partial_{-} \phi^1$ and $\partial_{+} \phi^2$ are modes that approaching the interface from their respective domain. We therefore establish the scattering relation 
\begin{eqnarray}\begin{aligned}
\label{eq:def_S}
\begin{pmatrix}
\partial_+\phi^1\\
\partial_-\phi^2
\end{pmatrix}
=S
\begin{pmatrix}
\partial_-\phi^1\\
\partial_+\phi^2
\end{pmatrix}
\end{aligned}\end{eqnarray}

\todo[inline]{Tianci: Add in the right panel in Fig.~\ref{fig:cut-and-join} the field content?}
and solve the $S$ matrix for the two cases of $M_1$ and $M_2$
\begin{eqnarray}\begin{aligned}
\label{eq:S1_S2}
S_1(\theta)=\begin{bmatrix}
\cos 2\theta & \sin 2\theta \\
\sin 2\theta & -\cos 2\theta
\end{bmatrix}\quad
S_2(\theta)=\begin{bmatrix}
-\cos 2\theta & \sin 2\theta \\
-\sin 2\theta & -\cos 2\theta
\end{bmatrix}
\end{aligned}\end{eqnarray}
\todo[inline]{Mao: please do not interchangably use defect or interface. We use interface throughout the text.}
For generic values of $\theta$, the interface is partially-transmitting, whose transmission and reflection coefficients can be read off from the scattering matrices. The fact of their wavelength independence is an indication of the interface's scale invariance.

We now work in the folding picture as shown in Fig.{\color{red} TODO}. The boundary at $x=0$ becomes impenetrable for the folded system, and the resulting tensor theory admits a conformal boundary state. The folding sends $\phi^2(x)$ to $\phi^2(-x)$ and hence the gluing condition becomes
\begin{eqnarray}\begin{aligned}
\partial_t(\cos\theta\phi^1-\sin\theta\phi^2)=0 \quad
\partial_x(\sin\theta\phi^1+\cos\theta\phi^2)=0 
\end{aligned}\end{eqnarray}
for the case $M=M_1(\theta)$.

If we quantize the Boson theory on the interface line $x = 0$, these gluing condition becomes an operator identity for the Boson creation and annihilation of the modes. We shall interpret these identities to be valid only when acting on the boundary states. The mode expansion of free Boson on $x = 0$\cite{di_francesco_conformal_1997} is
\begin{equation}
\phi(z, \bar{z} ) = \phi_0 - \frac{i}{4\pi g } \pi_0 \ln z \bar{z}  + \frac{i}{\sqrt{4\pi g} } \sum_{n \ne 0 } \frac{1}{n} \left(a_n z^n - \bar{a}_{-n } \bar{z}^{-n}   \right)
\end{equation}
where take $T$ to be time period and the following choice of the holomorphic and anti-holomorphic coordinates
\begin{equation}
z= e^{ 2\pi ( i\tau - x ) / T } \quad \bar{z} = e^{ -2\pi ( i\tau + x ) / T }
\end{equation}
We end up with a set of operator identities for each mode
\begin{eqnarray}\begin{aligned}
\label{eq:rotation_a_basis}
\cos\theta a_n^1-\sin\theta a_n^2 &= +( \cos\theta\bar{a}_{-n}^1-\sin\theta \bar{a}_{-n}^2 ) \\
\sin\theta a_n^1+\cos\theta a_n^2 &= -( \sin\theta\bar{a}_{-n}^1+\cos\theta \bar{a}_{-n}^2 ) 
\end{aligned}\end{eqnarray}
which is valid on the following boundary state
\begin{equation}
\label{eq:bd_state}
\begin{aligned}
| B \rangle 
& =  \exp\Big\{ \sum_{n > 0 } \frac{1}{n}
\begin{pmatrix}
a_{-n}^1\\
a_{-n}^2\\                              
\end{pmatrix}
S_1( \theta )
\begin{pmatrix}
\bar{a}_{-n}^1  \bar{a}_{-n}^2
\end{pmatrix} \Big\} |0\rangle
\end{aligned}
\end{equation}
where $S_1(\theta) $ is precisely the scattering matrix in Eq.~\eqref{eq:def_S}. The calculation for the case $M=M_2(\theta)$ is completely analogous. Eq.~\eqref{eq:bd_state} will be used extensively in the boundary state calculation of fidelity and Loschmidt echo in Sec.~\ref{sec:analytic_numerics}.

>>>>

\begin{equation}\begin{aligned}
\label{eq:def_M}
\begin{pmatrix}
\partial_t\phi^1\\
\partial_x\phi^1
\end{pmatrix}
=M\begin{pmatrix}
\partial_t\phi^2\\
\partial_x\phi^2
\end{pmatrix}
\end{aligned}\end{equation}
where derivative should be understood in the appropriate left and right limits, for example $\partial_x \phi^1$ is evaluated at $ x = 0^-$. As argued, the momentum component of stress tensor is continuous across the interface. As a consequence $M$ is an element of the Lorentz group $O(1,1)$ and can be parameterized as
\begin{equation}\begin{aligned}
M_1(\theta)=\pm
\begin{bmatrix}
\lambda^{-1} & 0 \\
0 & \lambda
\end{bmatrix}\quad
M_2(\theta)=\pm
\begin{bmatrix}
0 & \lambda  \\
\lambda^{-1} & 0 
\end{bmatrix}
\end{aligned}\end{equation}
where $\lambda=\tan\theta$ for $\theta\in\left[-\frac{\pi}{2},\frac{\pi}{2}\right]$.

Several special choice of $\theta$ need to be noted. 
\begin{enumerate}
\item $\theta=0,\pm \frac{\pi}{2}$. In this case, $\lambda$ (or $\lambda^{-1}$) appears to be singular and the fields on either side of the defect cannot transmit to the other side, so the interface reduces to individual boundary conditions for the boson on the left and right half planes: they are combination of Dirichlet and Neumann boundary conditions. For example, $\lambda = 0$ for $M_1$ implies $\partial_x\phi^1 = \partial_t\phi^2 =0$, which means that Dirichlet boundary condition are imposed on the right and Neumann boundary condition on the left. Hereafter we shall denote this combination as 'DN'. Similarly $M_1(\pm\frac{\pi}{2}),M_2(0),M_2(\pm \frac{\pi}{2})$ correspond to ND, DD, NN respectively. 
\item $\theta = \pm \frac{\pi}{4}$. In this case, $M(\theta)$ characterizes a perfectly transmitting defect. For example, there is effectively no defect in the case $M( \frac{\pi}{4})$ which is the traditional periodic boundary condition and we will denote it as P. For the other three cases, despite picking up a phase, the two counter propagating modes are still fully transmitted cross the interface. 
\end{enumerate}

The physical significance of $\theta$ will be manifest in the scattering process described below. We rewrite the Eq.~\eqref{eq:def_M} in the coordinates $t\pm x$ and $\partial_{\pm} = \partial_{t} \pm \partial_x $ derivative to extract the left and right going modes. For example, $\partial_{-} \phi^2$ will be a function $t - x$ only and hence represent a right going mode on the right half plane. It is then one of the scattering products that leave the interface. On the other hand $\partial_{-} \phi^1$ and $\partial_{+} \phi^2$ are modes that approaching the interface from their respective domain. We therefore establish the scattering relation 
\begin{equation}\begin{aligned}
\label{eq:def_S}
\begin{pmatrix}
\partial_+\phi^1\\
\partial_-\phi^2
\end{pmatrix}
=S
\begin{pmatrix}
\partial_-\phi^1\\
\partial_+\phi^2
\end{pmatrix}
\end{aligned}\end{equation}

and solve the $S$ matrix for the two cases of $M_1$ and $M_2$

\begin{equation}\begin{aligned}
\label{eq:S1_S2}
S_1(\theta)=\begin{bmatrix}
\cos 2\theta & \sin 2\theta \\
\sin 2\theta & -\cos 2\theta
\end{bmatrix}\quad
S_2(\theta)=\begin{bmatrix}
-\cos 2\theta & \sin 2\theta \\
-\sin 2\theta & -\cos 2\theta
\end{bmatrix}
\end{aligned}\end{equation}
\todo[inline]{Mao: please do not interchangably use defect or interface. We use interface throughout the text.}
For generic values of $\theta$, the interface is partially-transmitting, whose transmission and reflection coefficients can be read off from the scattering matrices. The fact of their wavelength independence is an indication of the interface's scale invariance.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{fig_folding_pic.pdf}
\caption{{\color{red} folding picture text}}
\label{fig:folding_pic}
\end{figure}

We now work in the folding picture as shown in Fig.~\ref{fig:folding_pic}. The boundary at $x=0$ becomes impenetrable for the folded system, and the resulting tensor theory admits a conformal boundary state. The folding sends $\phi^2(x)$ to $\phi^2(-x)$ and hence the gluing condition becomes
\begin{equation}
\begin{aligned}
\partial_t(\cos\theta\phi^1-\sin\theta\phi^2)=0 \quad
\partial_x(\sin\theta\phi^1+\cos\theta\phi^2)=0 
\end{aligned}
\end{equation}
for the case $M = M_1(\theta)$.

If we quantize the Boson theory on the interface line $x = 0$, these gluing condition becomes an operator identity for the Boson creation and annihilation of the modes. We shall interpret these identities to be valid only when acting on the boundary states. The mode expansion of free Boson on $x = 0$\cite{di_francesco_conformal_1997} is
\begin{equation}
\phi(z, \bar{z} ) = \phi_0 - \frac{i}{4\pi g } \pi_0 \ln z \bar{z}  + \frac{i}{\sqrt{4\pi g} } \sum_{n \ne 0 } \frac{1}{n} \left(a_n z^n - \bar{a}_{-n } \bar{z}^{-n}   \right)
\end{equation}
where take $T$ to be time period and the following choice of the holomorphic and anti-holomorphic coordinates
\begin{equation}
z= e^{ 2\pi ( i\tau - x ) / T } \quad \bar{z} = e^{ -2\pi ( i\tau + x ) / T }
\end{equation}
We end up with a set of operator identities for each mode
\begin{equation}
\begin{aligned}
\label{eq:rotation_a_basis}
\cos\theta a_n^1-\sin\theta a_n^2 &= +( \cos\theta\bar{a}_{-n}^1-\sin\theta \bar{a}_{-n}^2 ) \\
\sin\theta a_n^1+\cos\theta a_n^2 &= -( \sin\theta\bar{a}_{-n}^1+\cos\theta \bar{a}_{-n}^2 ) 
\end{aligned}
\end{equation}
which is valid on the following boundary state
\begin{equation}
\label{eq:bd_state}
\begin{aligned}
| B \rangle 
& =  \exp\Big\{ \sum_{n > 0 } \frac{1}{n}
\begin{pmatrix}
a_{-n}^1\\
a_{-n}^2\\                              
\end{pmatrix}
S_1( \theta )
\begin{pmatrix}
\bar{a}_{-n}^1  \bar{a}_{-n}^2
\end{pmatrix} \Big\} |0\rangle
\end{aligned}
\end{equation}
where $S_1(\theta) $ is precisely the scattering matrix in Eq.~\eqref{eq:def_S}. The calculation for the case $M=M_2(\theta)$ is completely analogous. Eq.~\eqref{eq:bd_state} will be used extensively in the boundary state calculation of fidelity and Loschmidt echo in Sec.~\ref{sec:analytic_numerics}.

### 2017-05-11 gnd_dn_lambda.tex
[draft] appendix done; analytic results done; discussion draft 1.

% flatex input: [gnd_dn_lambda.tex]

@@

In this appendix, we would like to calculate the amplitude for the setup shown in Fig.~\ref{fig:Fig_gnd_dn_lambda}. In particular, the unfolded configuration has D/N boundary conditions at $x = \pm \frac{L}{2}$ and linking boundary condition at $x = 0$. The general solutions can be written as
\begin{equation}
\label{eq:normalized_f}
f(k, x) = 
\left\lbrace
\begin{aligned}
  A_1 e^{i kt} \cos\left(kx +\frac{1}{2}kL \right) &  \quad x < 0  \\
  A_2 e^{ikt}  \sin\left(kx - \frac{1}{2}kL \right) & \quad x > 0   \\
\end{aligned} \right. 
\end{equation}
As demonstrated in Sec.~\ref{sec_sub:general_formulation}, if we denote $f(k,x<0)\equiv\phi_1$ and $f(k,x>0)\equiv\phi_2$, the boundary condition at the junction reads
\begin{eqnarray}\begin{aligned}
\frac{\partial_x \phi_1}{ \partial_t \phi_1} = \lambda^2 \frac{\partial_x \phi_2}{ \partial_t \phi_2} = \tan^2 \theta\frac{\partial_x \phi_2}{ \partial_t \phi_2}, \quad \theta \in \left[0,\frac{\pi}{2} \right]  
\end{aligned}\end{eqnarray}
which implies
\begin{equation}
\label{eq:momentum_gnd_dn_lambda}
k = \frac{2\pi}{L}\left( n \pm \frac{\theta}{\pi} \right),  \quad n\in\mathbb{Z}
\end{equation}
It is evident that the momentum $k$ is shifted from integer multiple of $2\pi/L$ due to the $\lambda$ boundary condition in the middle. 
\begin{comment}
Thus, the normalized eigenfunctions are
\begin{equation}
f_n(x) = \sqrt{\frac{2}{L}}
\left\lbrace
\begin{aligned}
  \cos(kx +\frac{1}{2}kL ) &  \quad x < 0  \\
  \pm \sin(kx - \frac{1}{2}kL ) & \quad x > 0   \\
\end{aligned} \right. 
\qquad 
k = \frac{2\pi}{L}\left( n \pm  \frac{\theta}{\pi} \right)  \quad n \in \mathbb{Z} 
\end{equation}
Expand the field $\phi = \sum_n \phi_n f_n(x) $, the action and Hamiltonian becomes
\begin{equation}
  S = \frac{g}{2} \int dt \, \sum_{n \in \mathbb{Z} }\left(  \dot{\phi}^2_n + k^2 \phi_n^2 \right) \implies\quad   g \dot{\phi}_n  = \pi_n \quad \implies H =  
\frac{1}{2g}\sum_{n \in \mathbb{Z} } \pi_n^2 + ( kg )^2  \phi_n^2 
\end{equation}
\end{comment}

It is clear that the normalized eigenfunctions in Eq.~\eqref{eq:normalized_f} serves as orthonormal basis in mode expansion. Following the same procedure in Ref.~\onlinecite{di_francesco_conformal_1997}, we have the Hamiltonian as
\begin{equation}
\label{eq:H_in_gnd_dn_lambda}
H = \frac{1}{2} \sum_{n \in \mathbb{Z} } |k|  \left(a^{\dagger}_n a_n + \frac{1}{2} \right)
\end{equation}
where the momentum $k$ is defined in Eq.~\eqref{eq:momentum_gnd_dn_lambda}, and the creation and annihilation operators defined as usual
\begin{equation}
\begin{aligned}
a_n = \frac{1}{\sqrt{2}} \left( \sqrt{ |k|g} \phi_n + \frac{i }{\sqrt{|k|g} }\pi_n  \right) \\
a^{\dagger}_n = \frac{1}{\sqrt{2}} \left( \sqrt{ |k|g} \phi_n - \frac{i }{\sqrt{|k|g} }\pi_n  \right) \\
\end{aligned}
\end{equation}

The Casimir energy is the vacuum energy brought by the finite size of the setup. From Eq.~\eqref{eq:H_in_gnd_dn_lambda} and define $x\equiv\theta/\pi$ in Eq.~\eqref{eq:momentum_gnd_dn_lambda}, we have
\begin{equation}
\begin{aligned}
E_c &= \frac{1}{4} \sum_{n \in \mathbb{Z}} | k| = \frac{\pi}{2L} \left( \sum_{n \in \mathbb{Z}}  | n + x | + \sum_{n \in \mathbb{Z}}  | n - x |  \right) \\  
&= \frac{\pi}{L} \Big[\sum_{n \ge 0 } ( n + x )^{-s} + \sum_{n \ge 0 }  ( n - x)^{-s}  +   x^{-s} \Big]\Big|_{s = -1} \\
&= \frac{\pi}{L} \left[ \zeta_{\rm H}( -1, x ) + \zeta_{\rm H}( -1, x ) +  x \right] \\
&= \frac{1}{2} \left( - x^2 + x - \frac{1}{6}\right)
\end{aligned}
\end{equation}
where we used the fact that the unfolded geometry has length $L=2\pi$. Thus the free energy reads
\begin{equation}
F = \beta E_c = - \frac{\beta}{2} B_2( x) 
\end{equation}
which agrees with the boundary state calculation in Sec.~\ref{app:lambda_12}.

>>>>

@@

In this appendix, we calculate the amplitude for the setup shown in Fig.~\ref{fig:Fig_gnd_dn_lambda}. In particular, the unfolded configuration has D/N boundary conditions at $x = \pm \frac{L}{2}$ and conformal interface $\lambda$ at $x = 0$. The general solutions can be written as
\begin{equation}
\label{eq:normalized_f}
f(k, x) = 
\left\lbrace
\begin{aligned}
  A_1 e^{i kt} \cos\left(kx +\frac{1}{2}kL \right) &  \quad x < 0  \\
  A_2 e^{ikt}  \sin\left(kx - \frac{1}{2}kL \right) & \quad x > 0   \\
\end{aligned} \right. 
\end{equation}
As demonstrated in Sec.~\ref{sec_sub:general_formulation}, if we denote $f(k,x<0)\equiv\phi_1$ and $f(k,x>0)\equiv\phi_2$, the boundary condition at the junction becomes
\begin{eqnarray}\begin{aligned}
\frac{\partial_x \phi_1}{ \partial_t \phi_1} = \lambda^2 \frac{\partial_x \phi_2}{ \partial_t \phi_2} = \tan^2 \theta\frac{\partial_x \phi_2}{ \partial_t \phi_2}, \quad \theta \in \left[0,\frac{\pi}{2} \right]  
\end{aligned}\end{eqnarray}
which implies
\begin{equation}
\label{eq:momentum_gnd_dn_lambda}
k = \frac{2\pi}{L}\left( n \pm \frac{\theta}{\pi} \right),  \quad n\in\mathbb{Z}
\end{equation}
It is evident that the momentum $k$ is shifted from integer multiple of $\frac{2\pi}{L}$ due to the $\lambda$ boundary condition in the middle.

The normalized eigenfunctions in Eq.~\eqref{eq:normalized_f} serves as orthonormal basis in the mode expansion, we thus have the 
\begin{equation}
\label{eq:H_in_gnd_dn_lambda}
H = \frac{1}{2} \sum_{n \in \mathbb{Z} } |k|  \left(a^{\dagger}_n a_n + \frac{1}{2} \right)
\end{equation}
where the momentum $k$ is defined in Eq.~\eqref{eq:momentum_gnd_dn_lambda}, and the creation and annihilation operators defined as usual
\begin{equation}
\begin{aligned}
a_n = \frac{1}{\sqrt{2}} \left( \sqrt{ |k|g} \phi_n + \frac{i }{\sqrt{|k|g} }\pi_n  \right) \\
a^{\dagger}_n = \frac{1}{\sqrt{2}} \left( \sqrt{ |k|g} \phi_n - \frac{i }{\sqrt{|k|g} }\pi_n  \right) \\
\end{aligned}
\end{equation}

The Casimir energy is the vacuum energy brought up by the finite size of the setup. From Eq.~\eqref{eq:H_in_gnd_dn_lambda} and use $x\equiv \frac{\theta}{\pi}$ in Eq.~\eqref{eq:momentum_gnd_dn_lambda}, we have
\begin{equation}
\begin{aligned}
E_c &= \frac{1}{4} \sum_{n \in \mathbb{Z}} | k| = \frac{\pi}{2L} \left( \sum_{n \in \mathbb{Z}}  | n + x | + \sum_{n \in \mathbb{Z}}  | n - x |  \right) \\  
&= \frac{\pi}{L} \Big[\sum_{n \ge 0 } ( n + x )^{-s} + \sum_{n \ge 0 }  ( n - x)^{-s}  +   x^{-s} \Big]\Big|_{s = -1} \\
&= \frac{\pi}{L} \left[ \zeta_{\rm H}( -1, x ) + \zeta_{\rm H}( -1, x ) +  x \right] \\
&= \frac{1}{2} \left( - x^2 + x - \frac{1}{6}\right)
\end{aligned}
\end{equation}
where we used the fact that the unfolded geometry has length $L=2\pi$. Thus the free energy in the large $\beta$ limit is
\begin{equation}
F = \beta E_c = - \frac{\beta}{2} B_2( x) 
\end{equation}
which agrees with the boundary state calculation in Sec.~\ref{app:lambda_12}.

%%% Local Variables:
%%% TeX-master: "bCFT_paper"
%%% TeX-PDF-mode: t
%%% End:

### 2017-05-11 intro.tex
[draft] appendix done; analytic results done; discussion draft 1.

% quantum impurity problem,  boundary critical phenomena, Anderson orthogonality catastrophe
In (one-dimensional) quantum critical systems, the presence of physical boundary and isolated impurity weakly break the conformal symmetry. Simply put, they scatter and hence relate the otherwise independent modes and therefore demonstrate novel boundary critical phenomena\cite{cardy_boundary_2004}. Operators close to the boundary are interpreted as boundary condition changing operator\cite{oshikawa_boundary_1997,affleck_boundary_1997} in the boundary conformal field theory (CFT). Their correlation functions exhibit different critical exponent from the bulk\cite{cardy_conformal_1984}. One example is the "Anderson orthogonality catastrophe", where the core hole creates a potential that act as an impurity to the conduction band. The X-ray absorption rate will then have a power singularity of boundary exponent\cite{affleck_boundary_1997} at the resonance frequency. There are numerous impurity problem of this kind that has been studied in the last few decades, such as magnetic impurity in spin chain\cite{eggert_magnetic_1992}, boundary and impurity effects in Luttinger liquid\cite{fabrizio_interacting_1995}, entanglement of the defects\cite{peschel_entanglement_2005, igloi_entanglement_2009,calabrese_entanglement_2012} \etc.

>>>>

% quantum impurity problem,  boundary critical phenomena, Anderson orthogonality catastrophe
In (one-dimensional) quantum critical systems, the presence of physical boundary and isolated impurity weakly break the conformal symmetry. Simply put, they scatter and hence relate the otherwise independent modes and therefore demonstrate novel boundary critical phenomena\cite{cardy_boundary_2004}. Operators close to the boundary are interpreted as boundary condition changing operator\cite{oshikawa_boundary_1997,affleck_boundary_1997} in the boundary conformal field theory (CFT). Their correlation functions exhibit different critical exponent from the bulk\cite{cardy_conformal_1984}. One example is the "Anderson orthogonality catastrophe", where the core hole creates a potential that act as an impurity to the conduction band. The X-ray absorption rate will then have a power singularity of boundary exponent\cite{affleck_boundary_1997} at the resonance frequency. There are numerous impurity problem of this kind that has been studied in the last few decades, such as magnetic impurity in spin chain\cite{eggert_magnetic_1992}, boundary and impurity effects in Luttinger liquid\cite{fabrizio_interacting_1995}, entanglement of the defects\cite{peschel_entanglement_2005, igloi_entanglement_2009,calabrese_entanglement_2012} \etc

### 2017-05-11 pf_of_id.tex
[draft] 1.add bibs, 2. geneneral formulation, free boson lattice done 3. Appendix of winding mode not done yet, I will add the caveat later
[draft] appendix done; analytic results done; discussion draft 1.

% flatex input: [pf_of_id.tex]

In this appendix, we provide more details for calculating the amplitude $Z_{ab}$ in Sec.{\bf\color{red}where}. We started to prove the following identity for a real symmetric matrix $M$
\begin{equation}
\label{eq:first_id_app_pf_of_id}
\exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{ \vec{b}^{\dagger} R \bar{\vec{b}}^{\dagger}  \Big\}  = \exp \Big\{ \vec{b}^{\dagger} e^{-M}  R \bar{\vec{b}}^{\dagger}  \Big\} \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} 
\end{equation}
where ${\bf b}$ and $\bar{\bf b}$ are vectors of bosonic operators. As emphasized in the {\bf\color{red}main text}, the matrix notation, for example, $\vec{b}^{\dagger} R \bar{\vec{b}}^{\dagger}$ should really mean $\sum_{ij}b^\dagger_iR_{ij}\bar{b}_j^\dagger$ where the dagger does not transpose the vector.

To prove Eq.~\eqref{eq:first_id_app_pf_of_id}, we first consider the special case where $R=\mathbb{I}$. We diagonalize $M = O^{T} \Lambda O $ and rotate the two sets of boson operators to the diagonal basis
\begin{equation}
  \vec{b}^{\dagger}  M \vec{b} = \vec{d}^{\dagger} \Lambda \vec{d}  \quad \vec{d} = O \vec{b} \quad \vec{\bar{d}}^\dagger = O^T \vec{\bar{b}}^\dagger
\end{equation}
where we understood $\vec{\bar{b}}^\dagger$ is a column vector and independent to $\vec{{b}}^\dagger$. Thus the whole expression can be written as
\begin{equation}
\begin{aligned}
  \exp&\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{ \vec{b}^{\dagger} \bar{\vec{b}}^\dagger  \Big\}  =  
  \exp\Big\{- \vec{d}^{\dagger} \Lambda \vec{d}  \Big\} \exp \Big\{   \vec{d}^{\dagger} \bar{\vec{d}}^\dagger  \Big\} \\
& = \prod_i  \exp\Big\{- \lambda_i d_i^{\dagger} d_i  \Big\} \exp \Big\{  d_i^{\dagger} \bar{d}_i ^{\dagger}  \Big\}
\end{aligned}
\end{equation}
We recall for $ [X, Y] = sY $, 
\begin{equation}
  e^X e^{Y} = e^{\exp (s ) Y} e^{X}
\end{equation}
which is a solvable case of Baker-Campbell-Hausdorff formula. Upon taking $X = -\lambda_i d_i^{\dagger} d_i$, $Y = d_i^{\dagger} \bar{d}^{\dagger}_i$, we have
\begin{equation}
\label{eq:lambda_commutator}
[- \lambda_i d_i^{\dagger} d_i, d_i ^{\dagger} \bar{d}_i^{\dagger}] =  - \lambda_i  d_i ^{\dagger} \bar{d}_i^{\dagger} 
\end{equation}
and so $s = - \lambda_i$ for each $\lambda_i$. This enable us to commute those exponentials
\begin{equation}
\begin{aligned}
 \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{  \vec{b}^{\dagger} \bar{\vec{b}}^\dagger  \Big\}   &= \prod_i \exp \Big\{ e^{- \lambda_i }  d^{\dagger}_i \bar{d}^{\dagger}_i  \Big\}  \exp \Big\{-\lambda_i  d^{\dagger}_i d_i  \Big\} \\
 & = \exp \Big\{ \vec{b}^{\dagger} e^{-M}  \bar{\vec{b}}^\dagger  \Big\} \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} 
\end{aligned}
\end{equation}
For the general case where $R \neq\mathbb{I}$, we take $\vec{\bar{d}}^* = O^T R \vec{\bar{b}}^*$. This will not change the commutation relation of ${\bf d}$, and the role of $\bar{b}$ is decorative in Eq.~\eqref{eq:lambda_commutator}. Hence the rest of the proof follows the same way. \hfill$\blacksquare$

>>>>

In this appendix, we provide more details for calculating the amplitude $Z_{ab}$ in Sec.~\ref{sec:analytic_numerics}. We start to prove the following identity for a real symmetric matrix $M$
\begin{equation}
\label{eq:first_id_app_pf_of_id}
\exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{ \vec{b}^{\dagger} R \bar{\vec{b}}^{\dagger}  \Big\}  = \exp \Big\{ \vec{b}^{\dagger} e^{-M}  R \bar{\vec{b}}^{\dagger}  \Big\} \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} 
\end{equation}
where ${\bf b}$ and $\bar{\bf b}$ are vectors of bosonic operators. The matrix notation here should be understand as bilinear form. For example, $\vec{b}^{\dagger} R \bar{\vec{b}}^{\dagger}$ should really mean $\sum_{ij}b^\dagger_iR_{ij}\bar{b}_j^\dagger$ where the dagger does not transpose the vector.

To prove Eq.~\eqref{eq:first_id_app_pf_of_id}, we first consider the special case where $R=\mathbb{I}$. We diagonalize $M = O^{T} \Lambda O $ and rotate the two sets of boson operators to the diagonal basis
\begin{equation}
  \vec{b}^{\dagger}  M \vec{b} = \vec{d}^{\dagger} \Lambda \vec{d}  \quad \vec{d} = O \vec{b} \quad \vec{\bar{d}}^\dagger = O^T \vec{\bar{b}}^\dagger
\end{equation}
where we understand $\vec{\bar{b}}^\dagger$ as a column vector independent to $\vec{{b}}^\dagger$. Thus the whole expression can be written as
\begin{equation}
\begin{aligned}
  \exp&\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{ \vec{b}^{\dagger} \bar{\vec{b}}^\dagger  \Big\}  =  
  \exp\Big\{- \vec{d}^{\dagger} \Lambda \vec{d}  \Big\} \exp \Big\{   \vec{d}^{\dagger} \bar{\vec{d}}^\dagger  \Big\} \\
& = \prod_i  \exp\Big\{- \lambda_i d_i^{\dagger} d_i  \Big\} \exp \Big\{  d_i^{\dagger} \bar{d}_i ^{\dagger}  \Big\}
\end{aligned}
\end{equation}
We recall for $ [X, Y] = sY $, 
\begin{equation}
  e^X e^{Y} = e^{\exp (s ) Y} e^{X}
\end{equation}
which is a solvable case of Baker-Campbell-Hausdorff formula. Upon taking $X = -\lambda_i d_i^{\dagger} d_i$, $Y = d_i^{\dagger} \bar{d}^{\dagger}_i$, we have
\begin{equation}
\label{eq:lambda_commutator}
[- \lambda_i d_i^{\dagger} d_i, d_i ^{\dagger} \bar{d}_i^{\dagger}] =  - \lambda_i  d_i ^{\dagger} \bar{d}_i^{\dagger} 
\end{equation}
and so $s = - \lambda_i$ for each $\lambda_i$. This enable us to commute those exponentials
\begin{equation}
\begin{aligned}
 \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} \exp \Big\{  \vec{b}^{\dagger} \bar{\vec{b}}^\dagger  \Big\}   &= \prod_i \exp \Big\{ e^{- \lambda_i }  d^{\dagger}_i \bar{d}^{\dagger}_i  \Big\}  \exp \Big\{-\lambda_i  d^{\dagger}_i d_i  \Big\} \\
 & = \exp \Big\{ \vec{b}^{\dagger} e^{-M}  \bar{\vec{b}}^\dagger  \Big\} \exp\Big\{- \vec{b}^{\dagger} M \vec{b}  \Big\} 
\end{aligned}
\end{equation}
For the general case where $R \neq\mathbb{I}$, we take $\vec{\bar{d}}^* = O^T R \vec{\bar{b}}^*$. This will not change the commutation relation of ${\bf d}$, and the role of $\bar{b}$ is decorative in Eq.~\eqref{eq:lambda_commutator}. Hence the rest of the proof follows the same way. \hfill$\blacksquare$
